### 2023-06-27 manuscript/supplement.qmd
started a supplement file

---
title: "Supplementary Appendix"
subtitle: "The FAIRification of pharmacoepidemiological research: A practical introduction to reproducible analytical workflows using git and R"
code-fold: true
eval: false
echo: true
number-sections: true

format: 
  pdf:
    fig-cap-location: top
    crossref:
    fig-title: "**Supplementary Figure**"

editor: visual
bibliography: references.bib
csl: pharmacoepidemiology-and-drug-safety.csl
---

```{r setup}
#| label: setup
#| include: false
#| eval: true
library(magick)
library(cowplot)
library(here)
```

**Authors**: Janick Weberpals^1^, Shirley V. Wang^1^

**Author affiliation**: ^1^Division of Pharmacoepidemiologyand Pharmacoeconomics, Department of Medicine, Brigham and Women's Hospital, Harvard Medical School, Boston, MA

# Supplementary Figures

```{r}
#| label: fig-GitHub-gui_1
#| fig-cap: "Starting a new repository through the GitHub GUI."
#| eval: false
#| echo: false
#| out-width: 60%

sfig_a <- ggdraw() + 
  draw_image(image_read(here("manuscript/figures", "Figure_1a_new_repo.png")), scale = 0.5)

sfig_a
```

{{< pagebreak >}}

```{r}
#| label: fig-GitHub-gui
#| fig-cap: ""
#| eval: false
#| echo: false
#| out-width: 60%

sfig_a <- ggdraw() + 
  draw_image(image_read(here("manuscript/figures", "Figure_1a_new_repo.png")), scale = 0.5)

sfig_b <- ggdraw() + 
  draw_image(image_read(here("manuscript/figures", "Figure_1b_repo_info.png")))

sfig_c <- ggdraw() + 
  draw_image(image_read(here("manuscript/figures", "Figure_1c_clone.png")))

sfig1 <- plot_grid(
  fig_a, fig_c, fig_b, 
  nrow = 2,
  ncol = 2
)

sfig1
```

{{< pagebreak >}}

# References {.unnumbered}