### 2023-06-27 manuscript/supplement.qmd
started a supplementary document

---
title: "Supplementary Appendix"
subtitle: "The FAIRification of pharmacoepidemiological research: A practical introduction to reproducible analytical workflows using git and R"
code-fold: true
echo: false
number-sections: true

format: 
  pdf:
    fig-cap-location: top
    fig-align: center
    crossref:
      fig-title: "**Supplementary Figure**"
    toc: true
    toc-depth: 2
    number-sections: false
    colorlinks: true

editor: visual
bibliography: references.bib
csl: pharmacoepidemiology-and-drug-safety.csl
---

```{r setup}
#| label: setup
#| include: false
#| eval: true
library(magick)
library(cowplot)
library(here)
library(knitr)
```

{{< pagebreak >}}

# Supplementary Figures

```{r}
#| label: fig-GitHub-gui_a
#| fig-cap: "Starting a new repository through the GitHub GUI."
#| out-width: 115%

include_graphics(here("manuscript/figures", "SFigure_1a_new_repo.png"))
```

```{r}
#| label: fig-GitHub-gui_b
#| fig-cap: "Populating basic information about the new repository."
#| out-width: 110%

include_graphics(here("manuscript/figures", "SFigure_2_repo_info.png"))
```

```{r}
#| label: fig-GitHub-gui_c
#| fig-cap: "Clone a GitHub repository using the HTTPS link and the git clone command"
#| out-width: 115%

include_graphics(here("manuscript/figures", "SFigure_3_clone.png"))
```

{{< pagebreak >}}

# References {.unnumbered}